### 2019-10-22 Manuscript/Manuscript.tex
fixing overleaf

Over geologic time-scales, \COtwo enters Earth's ocean-atmosphere system primarily via volcanism and metamorphic degassing, and leaves primarily through the chemical weathering of silicate rocks, which delivers alkalinity and cations to the ocean that sequester carbon via the precipitation of carbonate sediment, with a smaller sink through organic carbon burial \cite{Kump2000a}. Steady-state \pCOtwo is set at the \pCOtwo level at which \COtwo sinks are equal to sources. As \COtwo sinks are removed and \pCOtwo rises, temperature increases and the hydrological cycle is invigorated, causing other weathering sinks to increase until a new steady-state is achieved at higher \pCOtwo. Conversely, the magnitude of the \COtwo sink can be equal to the magnitude of the source at lower \pCOtwo when global weatherability is high \cite{Kump1997a}.

Topography, climate, and lithology all modulate chemical weathering. In contrast to regions of high denudation, in low-relief regions low erosion rates lead to supply-limited weathering regimes \cite{Gabet2009a, West2012a, Maher2014a}. In warm and wet regions, mineral dissolution occurs more readily leading to enhanced chemical weathering \cite{Lasaga1994a, West2012a}. Regions with high physical weathering rates that are not supply-limited are more sensitive to changes in climate that can reduce or enhance chemical weathering \cite{West2012a, Maher2014a}. In terms of lithology, mafic rocks have the potential to sequester more carbon through silicate weathering because of their relatively high Ca and Mg concentrations \cite{Dessert2003a}. These factors have lead to the proposal that arc-continent collisions within the tropical rain belt have been important in enhancing global weatherability, lowering atmospheric \pCOtwo, and initiating glacial climate over the past 520~m.y. \cite{Jagoutz2016a, Swanson-Hysell2017a, Macdonald2019a} and perhaps in the Neoproterozoic as well \cite{Park2019a}.

@@

The silicate weathering component of GEOCLIM calculates \COtwo consumption resulting from silicate weathering for each continental grid cell. In previous versions of the model, silicate weathering was a function of temperature and runoff only, and bedrock in all continental grid cells had identical chemical compositions \cite{Godderis2014a}. More recent versions of GEOCLIM implement regolith development and soil shielding into the silicate weathering component, which introduces a dependence on erosion (and then topographic slope see SI) \cite{Gabet2009a, West2012a}. While this introduction of regolith development into GEOCLIM is important for assessing the impact of tropical arc-continent collisions on \pCOtwo, the relatively high Ca+Mg concentration in arc rocks relative to other lithologies must also be considered. We therefore implement variable bedrock Ca+Mg concentration into GEOCLIM. The spatial distribution of lithologies is sourced from the Global Lithologic Map (GLiM) \cite{Hartmann2012a} and is represented by 6 categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment (see SI). Each continental grid cell is assigned these lithologic categories at a resolution of 0.1\degrees $\times$ 0.1\degrees. The Ca+Mg concentrations of felsic, intermediate, and mafic lithologies are assigned based on the mean of data compiled from EarthChem (\url{www.earthchem.org/portal}). Given that GLiM does not distinguish ultramafic lithologies, such rocks are grouped with mafic rocks. As a result, the Ca+Mg concentration is likely an underestimate in regions of obducted ophiolites, such that the estimated effect of these regions on changing steady-state \pCOtwo is conservative. The weathering of carbonate does not contribute to long-term \COtwo consumption and its Ca+Mg concentration is ignored. The Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are more difficult to define, since their chemical composition is strongly dependent on protolith composition and, in the case of siliciclastic sediment, the degree of previous chemical depletion. We explore a range of feasible Ca+Mg concentrations of metamorphic and siliciclastic sediment lithologies during calibration of the silicate weathering component.

@@

Having calibrated the silicate weathering component of GEOCLIM, we use it to estimate the decrease in steady-state \pCOtwo associated with the emergence of SEAI. For the climate model component, we use temperature and runoff from a subset of the GFDL CM2.0 experiments \cite{Delworth2006a, Delworth2006b} (see SI). The primary strength of these experiments for this analysis is that all non-\COtwo forcings are held constant at values representative of pre-industrial conditions, allowing the effect of \pCOtwo to be isolated. We use terrestrial and marine sedimentary deposits to determine the position of paleoshorelines of SEAI over the past 15~m.y. and include changes in areas of presently-submerged continental shelves such as the Sunda Shelf that were previously exposed (Fig. \ref{fig:shoreline_growth}; see SI). In New Guinea, emergence in the Middle Miocene is associated with collision between the Melanesian Arc and Australia's distal margin \cite{vanUfford2005a, Cloos2005a, Baldwin2012a}, which drove exhumation of the 1440~km long Irian-Marum-April Ophiolite Belt. Exhumation of the New Guinea Central Range accelerated over the past 4~m.y. associated with slab-breakoff and buoyant uplift \cite{Cloos2005a} while in eastern New Guinea, progressive jamming of the north-dipping subduction zone also caused major uplift over this time interval \cite{vanUfford2005a}. These tectonic drivers and other throughout the region lead to progressive emergence over the past 15~m.y. that accelerated following 5~Ma (Fig. \ref{fig:shoreline_growth}B). This trend mirrors broad cooling over the Neogene that resulted in the initiation of Northern Hemisphere ice sheets (Fig. \ref{fig:shoreline_growth}C).

We use GEOCLIM to estimate \pCOtwo associated with the reconstructed subaerial extent of islands in the area at ca. 15, 10, and 5~Ma (``reduced SEAI + Sunda Shelf'' scenarios; Fig. \ref{fig:scenario_pCO2}). Given that the Sunda Shelf is presently submerged, across the shelf we distribute felsic and siliciclastic lithologies and assign shallow slope values in a manner that is consistent with existing observations, we also extrapolate the runoff field from surrounding lands (see SI). While we modify the land area, the position of the tectonic blocks remain fixed since we use a climate model forced with modern geography. While there has been motion of these tectonic blocks since 15~Ma, they have remianed within tropical latitudes such that this fixed scenario is a good approximation of the paleogeography (see SI). We also test an end-member scenario, in which all islands associated with arc-continent collision in the region are removed (``removed SEAI'' scenario; Fig. \ref{fig:scenario_pCO2}).

Using the 3,339 unique parameter combinations, the ``reduced SEAI + Sunda Shelf'' scenarios resulted in 533--611~ppm, 466--502~ppm, and 394--412~ppm \pCOtwo for 15, 10, and 5~Ma respectively (Fig. \ref{fig:scenario_pCO2}). These results indicate a progressive decrease in \pCOtwo over the Neogene associated with the emergence of SEAI, and suggest that without this emergence, pre-industrial \pCOtwo would have been $\sim$533--611~ppm. Proxy-based estimates of the magnitude and trajectory of \pCOtwo change from the Miocene to the Pliocene are variable between techniques and associated assumptions underlying their interpretation \cite{Foster2017a, Mejia2017a, Ji2018a}. The modeled \pCOtwo values for 15~Ma resemble the higher end of proxy-based \pCOtwo estimates for the early-mid-Miocene (see SI), indicating that the increase in subaerially-exposed land area and tectonic topography of SEAI is sufficient to explain long-term cooling of Earth's climate over the Neogene. The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750~ppm with that for Northern Hemisphere glaciation being significantly lower at $\sim$280~ppm \cite{DeConto2008a}. These modeled values of decreasing \pCOtwo are therefore consistent with the record of Neogene climate with Miocene ice sheets on Antarctica followed by Northern Hemisphere ice sheets developing in the Pliocene as \pCOtwo subsequently decreased.

>>>>

Over geologic time-scales, \COtwo enters Earth's ocean--atmosphere system primarily via volcanism and metamorphic degassing, and leaves primarily through the chemical weathering of silicate rocks, which delivers alkalinity and cations to the ocean that sequester carbon via the precipitation of carbonate sediment, with a smaller sink through organic carbon burial \cite{Kump2000a}. Steady-state \pCOtwo is set at the \pCOtwo level at which \COtwo sinks are equal to sources. As \COtwo sinks are removed and \pCOtwo rises, temperature increases and the hydrological cycle is invigorated, causing other weathering sinks to increase until a new steady-state is achieved at higher \pCOtwo. Conversely, the magnitude of the \COtwo sink can be equal to the magnitude of the source at lower \pCOtwo when global weatherability is high \cite{Kump1997a}.

Topography, climate, and lithology all modulate chemical weathering. In contrast to regions of high denudation, in low-relief regions low erosion rates lead to supply-limited weathering regimes \cite{Gabet2009a, West2012a, Maher2014a}. In warm and wet regions, mineral dissolution occurs more readily leading to enhanced chemical weathering \cite{Lasaga1994a, West2012a}. Regions with high physical weathering rates that are not supply-limited are more sensitive to changes in climate that can reduce or enhance chemical weathering \cite{West2012a, Maher2014a}. In terms of lithology, mafic rocks have the potential to sequester more carbon through silicate weathering because of their relatively high Ca and Mg concentrations \cite{Dessert2003a}. These factors have lead to the proposal that arc-continent collisions within the tropical rain belt have been important in enhancing global weatherability, lowering atmospheric \pCOtwo, and initiating glacial climate over the past 520~m.y. \cite{Jagoutz2016a, Swanson-Hysell2017a, Macdonald2019a} and perhaps in the Neoproterozoic as well \cite{Park2019b}.

@@

The silicate weathering component of GEOCLIM calculates \COtwo consumption resulting from silicate weathering for each continental grid cell. In previous versions of the model, silicate weathering was a function of temperature and runoff only, and bedrock in all continental grid cells had identical chemical compositions \cite{Godderis2014a}. More recent versions of GEOCLIM implement regolith development and soil shielding into the silicate weathering component, which introduces a dependence on erosion (and then topographic slope, see SI) \cite{Gabet2009a, West2012a}. While this introduction of regolith development into GEOCLIM is important for assessing the impact of tropical arc-continent collisions on \pCOtwo, the relatively high Ca+Mg concentration in arc rocks relative to other lithologies must also be considered. We therefore implement variable bedrock Ca+Mg concentration into GEOCLIM. The spatial distribution of lithologies is sourced from the Global Lithologic Map (GLiM) \cite{Hartmann2012a} and is represented by 6 categories: metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment (see SI). Each continental grid cell is assigned these lithologic categories at a resolution of 0.1\degrees $\times$ 0.1\degrees. The Ca+Mg concentrations of felsic, intermediate, and mafic lithologies are assigned based on the mean of data compiled from EarthChem (\url{www.earthchem.org/portal}). Given that GLiM does not distinguish ultramafic lithologies, such rocks are grouped with mafic rocks. As a result, the Ca+Mg concentration is likely an underestimate in regions of obducted ophiolites, such that the estimated effect of these regions on changing steady-state \pCOtwo is conservative. The weathering of carbonate does not contribute to long-term \COtwo consumption and its Ca+Mg concentration is ignored. The Ca+Mg concentrations of metamorphic and siliciclastic sediment grid cells are more difficult to define, since their chemical composition is strongly dependent on protolith composition and, in the case of siliciclastic sediment, the degree of previous chemical depletion. We explore a range of feasible Ca+Mg concentrations of metamorphic and siliciclastic sediment lithologies during calibration of the silicate weathering component.

@@

Having calibrated the silicate weathering component of GEOCLIM, we use it to estimate the decrease in steady-state \pCOtwo associated with the emergence of SEAI. For the climate model component, we use temperature and runoff from a subset of the GFDL CM2.0 experiments \cite{Delworth2006b} (see SI). The primary strength of these experiments for this analysis is that all non-\COtwo forcings are held constant at values representative of pre-industrial conditions, allowing the effect of \pCOtwo to be isolated. We use terrestrial and marine sedimentary deposits to determine the position of paleoshorelines of SEAI over the past 15~m.y. and include changes in areas of presently-submerged continental shelves such as the Sunda Shelf that were previously exposed (Fig. \ref{fig:shoreline_growth}; see SI). In New Guinea, emergence in the Middle Miocene is associated with collision between the Melanesian Arc and Australia's distal margin \cite{vanUfford2005a, Cloos2005a, Baldwin2012a}, which drove exhumation of the 1440~km long Irian-Marum-April Ophiolite Belt. Exhumation of the New Guinea Central Range accelerated over the past 4~m.y. associated with slab-breakoff and buoyant uplift \cite{Cloos2005a} while in eastern New Guinea, progressive jamming of the north-dipping subduction zone also caused major uplift over this time interval \cite{vanUfford2005a}. These tectonic drivers and other throughout the region lead to progressive emergence over the past 15~m.y. that accelerated following 5~Ma (Fig. \ref{fig:shoreline_growth}B). This trend mirrors broad cooling over the Neogene that resulted in the initiation of Northern Hemisphere ice sheets (Fig. \ref{fig:shoreline_growth}C).

We use GEOCLIM to estimate \pCOtwo associated with the reconstructed subaerial extent of islands in the area at ca. 15, 10, and 5~Ma (``reduced SEAI + Sunda Shelf'' scenarios; Fig. \ref{fig:scenario_pCO2}). Given that the Sunda Shelf is presently submerged, across the shelf we distribute felsic and siliciclastic lithologies and assign shallow slope values in a manner that is consistent with existing observations, and extrapolate runoff from surrounding lands (see SI). While we modify the land area, the position of the tectonic blocks remain fixed since we use a climate model forced with modern geography. While there has been motion of these tectonic blocks since 15~Ma, they have remained within tropical latitudes such that this fixed scenario is a good approximation of the paleogeography (see SI). We also test an end-member scenario, in which all islands associated with arc-continent collision in the region are removed (``removed SEAI'' scenario; Fig. \ref{fig:scenario_pCO2}).

Using the 3,339 unique parameter combinations, the ``reduced SEAI + Sunda Shelf'' scenarios resulted in 533--611~ppm, 466--502~ppm, and 394--412~ppm \pCOtwo for 15, 10, and 5~Ma respectively (Fig. \ref{fig:scenario_pCO2}). These results indicate a progressive decrease in \pCOtwo over the Neogene associated with the emergence of SEAI, and suggest that without this emergence, pre-industrial \pCOtwo would have been $\sim$533--611~ppm. Proxy-based estimates of the magnitude and trajectory of \pCOtwo change from the Miocene to the Pliocene are variable between techniques and associated assumptions underlying their interpretation \cite{Foster2017a, Mejia2017a, Ji2018a}. The modeled \pCOtwo values for 15~Ma resemble the higher end of proxy-based \pCOtwo estimates for the early-mid-Miocene (see SI), indicating that the increase in subaerially-exposed land area and tectonic topography of SEAI is sufficient to explain long-term cooling of Earth's climate over the Neogene. The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750~ppm with that for Northern Hemisphere glaciation being significantly lower at $\sim$280~ppm \cite{DeConto2008a}. These modeled values of decreasing \pCOtwo are therefore consistent with the record of Neogene climate with Miocene ice sheets on Antarctica \cite{Sugden1995a} followed by Northern Hemisphere ice sheets developing in the Pliocene \cite{Haug2005a} as \pCOtwo subsequently decreased.

### 2019-10-22 Manuscript/Supplementary_Information/Supplementary_Information.tex
fixing overleaf

\noindent
$k_{e}$ is a proportionality constant ((m/yr)$^{1-m}$) and $m$ and $n$ are adjustable exponents that are kept as 0.5 and 1 as in \citet{Maffre2018a}. This formulation is directly inspired from the stream power law \citep{Davy2000a}. This law and these exponent values are supported by various compilations (e.g. \citealp{Lague2013a}) although \citet{Lague2013a} also suggests that there is variability in the proportionality constant that is difficult to capture at a global scale.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{Manuscript/Figures/weathering_regolith_maps.jpg}
    \caption{Chemical weathering and regolith thickness maps associated with the GEOCLIM model run using the parameter combination that produced that highest $r^{2}$ (Fig. \ref{fig:r2_cross_plot}B). Parameter values used for this model run are: $k_{d}=5\times10^{-4}$, $k_{w}=1$, $\sigma=-0.4$, $k_{rp}=1\times10^{-2}$, metamorphic$_{Ca+Mg}=2500$~mol/m$^{3}$, sediment$_{Ca+Mg}=2000$~mol/m$^{3}$.}
    \label{fig:weathering_regolith_maps}
\end{figure}

@@

We then compute spatially-resolved long-term \COtwo consumption (i.e. Ca+Mg fluxes) using present-day runoff (UNH/GRDC Composite Runoff Fields V1.0; \citealp{Fekete1999a}), temperature (CRU TS v.4.03; \citealp{Harris2013a}), and slope (Shuttle Radar Topography Mission; \citealp{Farr2007a}) fields. As described in the main text, we then sum the computed \COtwo consumption over large-scale watersheds that appear in the global compilation of \citet{Gaillardet1999a}, as well as smaller-scale watersheds of the Amazon Basin (HYBAM network) in the compilation of \citet{Moquet2011a, Moquet2016a, Moquet2018a}. Those lasts are nested watersheds, which means we have to subtract upstream weathering fluxes from downstream fluxes. Watersheds for which this subtraction yields an aberrant value are not considered. 80 watersheds in total are used in this study. We calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these basins:

@@

Global CO$_{2}$ consumption calculated from these 3,339 parameter combinations ranges from 2.7$\times10^{12}$~mol/yr to 7.1$\times10^{12}$~mol/yr, with a mean of 5.4$\times10^{12}$~mol/yr. These estimates fall within the range of those independently estimated in the literature for the present-day: the full range of estimates of the present-day non-anthropogenic global CO$_{2}$ emission rate compiled in \citet{Gerlach2011a} is 3.0--10.0$\times10^{12}$~mol/yr, with the preferred range of those estimates being 3.4--5.9$\times10^{12}$~mol/yr (Fig. \ref{fig:W_vs_r2}B).
Each parameter combination predicts a different total \COtwo consumption that should match the \COtwo degassing at steady-state. Hence, in the GEOCLIM experiments, each parameter combination is used with its corresponding steady-state \COtwo degassing. In these estimates we are not including the effects of reverse weathering \citep{Michalopoulos1995a} as it is not clear how it should be parameterized and it is interpreted to be a relatively minor flux in the Cenozoic \citep{Isson2018a}. Formation of authigenic clays that scales with Ca and Mg concentration of riverine waters could decrease the \COtwo consumption associated with silicate weathering although the overall trend seen in this study would remain.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{Manuscript/Figures/scenario_pCO2_all.pdf}
    \caption{Steady-state \pCOtwo estimates from GEOCLIM for the various scenarios discussed in the text. For each of the ten scenarios, each point represents an estimate from one of the 3,339 unique parameter combinations that resulted in reasonable total global \COtwo consumption and most closely matched \COtwo consumption estimated from watershed data (see SI). The box encloses the middle 50\% of the \pCOtwo estimates (i.e. the interquartile range), and the notch represents the median with its 95\% confidence interval. The whiskers extend to 1.5$\times$ the interquartile range, or to the minimum/maximum of the estimates if no estimates are below/above this range.}
    \label{fig:scenario_pCO2_all}
\end{figure}

We use geological data to quantify the changes in the area of I\&NG over the past 20~m.y. (described in detail in \textit{Paleoshoreline Reconstruction and Geological Synthesis}). To generate the lithologic map used in the ``removed I\&NG'' scenario (Fig. \ref{fig:ING_fracs}), we remove all land associated with arc-continent collision in I\&NG. To generate the lithologic map used in the ``reduced I\&NG + Sunda Shelf'' scenarios (Fig. \ref{fig:ING_fracs}), we take the estimated paleoshorelines of I\&NG for 15, 10, and 5~Ma, then remove all land that falls outside of these bounds. The slope and lithologic classification of each pixel is identical to that of the present day, provided that these pixels fall within the estimated paleoshorelines of I\&NG for their respective time slice. This means that while we are changing the amount of emergent land, we are not changing the latitudinal or longitudinal position and are keeping the slope of emerged pixels the same. Given that similar distributions of lithology and slope are reasonable in the past and that latitudinal translation has been relatively minor, we consider this simplification to be reasonable to first-order. We then add land that is not exposed today (i.e. the Sunda Shelf). However, the assignment of lithology, runoff, and slope for pixels on the Sunda Shelf is not trivial, given that the shelf is not currently exposed. As discussed in \textit{Paleoshoreline Reconstruction and Geological Synthesis}, the Sunda Shelf is a flat-lying, relatively stable platform of continental crust. Islands between Malaysia/Sumatra and Borneo are composed of granite and are surrounded by shallow marine siliciclastic sediments \citep{Darmadi2007a, Hall2009a, Hall2013b}. However, basement highs do not appear in the Gulf of Thailand, and seismic and drill core data suggest siliciclastic fluvial systems draining out toward the South China Sea when the shelf was exposed \citep{Darmadi2007a}. We therefore assign pixels of the Sunda Shelf between Malaysia/Sumatra and Borneo to be 50\% felsic and 50\% sediment, and pixels further to the north to be 100\% sediment. For the slope, we take the mean slope value (0.0038~m/m) of relatively flat land on eastern Sumatra, and assign this value to all pixels of the Sunda Shelf.  For the runoff, we linearly interpolate along latitude bands between present-day land pixels to obtain runoff values for the presently-submerged Sunda Shelf. This approach preserves large N-S variation in runoff associated with the large-scale Hadley circulation and therefore should be reasonable to first-order, although it does not accurately capture the smaller-scale interactions between land and ocean/atmosphere circulation.

>>>>

\noindent
$k_{e}$ is a proportionality constant ((m/yr)$^{1-m}$) and $m$ and $n$ are adjustable exponents that are kept as 0.5 and 1 as in \citet{Maffre2018a}. This formulation is directly inspired by the stream power law \citep{Davy2000a}. This formulation and these exponent values are supported by various compilations (e.g. \citealp{Lague2013a}) although \citet{Lague2013a} also suggests that there is variability in the proportionality constant that is difficult to capture at a global scale.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{Manuscript/Figures/weathering_regolith_maps.jpg}
    \caption{Chemical weathering and regolith thickness maps associated with the GEOCLIM model run using the parameter combination that produced the highest $r^{2}$ (Fig. \ref{fig:r2_cross_plot}B). Parameter values used for this model run are: $k_{d}=5\times10^{-4}$, $k_{w}=1$, $\sigma=-0.4$, $k_{rp}=1\times10^{-2}$, metamorphic$_{Ca+Mg}=2500$~mol/m$^{3}$, sediment$_{Ca+Mg}=2000$~mol/m$^{3}$.}
    \label{fig:weathering_regolith_maps}
\end{figure}

@@

We then compute spatially-resolved long-term \COtwo consumption (i.e. Ca+Mg fluxes) using present-day runoff (UNH/GRDC Composite Runoff Fields V1.0; \citealp{Fekete1999a}), temperature (CRU TS v.4.03; \citealp{Harris2013a}), and slope (Shuttle Radar Topography Mission; \citealp{Farr2007a}) fields. As described in the main text, we then sum the computed \COtwo consumption over large-scale watersheds that appear in the global compilation of \citet{Gaillardet1999a}, as well as smaller-scale watersheds of the Amazon Basin (HYBAM network) in the compilation of \citet{Moquet2011a, Moquet2016a, Moquet2018a}. The latter are nested watersheds, which requires upstream weathering fluxes to be subtracted from downstream fluxes. Watersheds for which this subtraction yields an aberrant value are not considered. 80 watersheds in total are used in this study. We calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these basins:

@@

Global CO$_{2}$ consumption calculated from these 3,339 parameter combinations ranges from 2.7$\times10^{12}$~mol/yr to 7.1$\times10^{12}$~mol/yr, with a mean of 5.4$\times10^{12}$~mol/yr. These estimates fall within the range of those independently estimated in the literature for the present-day: the full range of estimates of the present-day non-anthropogenic global CO$_{2}$ emission rate compiled in \citet{Gerlach2011a} is 3.0--10.0$\times10^{12}$~mol/yr, with the preferred range of those estimates being 3.4--5.9$\times10^{12}$~mol/yr (Fig. \ref{fig:W_vs_r2}B).

Each parameter combination predicts a different total \COtwo consumption that should match the \COtwo degassing at steady-state. Hence, in the GEOCLIM experiments, each parameter combination is used with its corresponding steady-state \COtwo degassing. In these estimates we are not including the effects of reverse weathering \citep{Michalopoulos1995a} as it is not clear how it should be parameterized and it is interpreted to be a relatively minor flux in the Cenozoic \citep{Isson2018a}. Formation of authigenic clays that scales with Ca and Mg concentration of riverine waters could decrease the \COtwo consumption associated with silicate weathering although the overall trend seen in this study would remain.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{Manuscript/Figures/scenario_pCO2_all.pdf}
    \caption{Steady-state \pCOtwo estimates from GEOCLIM for the various scenarios discussed in the text. For each of the ten scenarios, each point represents an estimate from one of the 3,339 unique parameter combinations that resulted in reasonable total global \COtwo consumption and most closely matched \COtwo consumption estimated from watershed data (see SI). The box encloses the middle 50\% of the \pCOtwo estimates (i.e. the interquartile range), and the notch represents the median with its 95\% confidence interval. The whiskers extend to 1.5$\times$ the interquartile range, or to the minimum/maximum of the estimates if no estimates are below/above this range. The only difference with Fig. X in the main text....}
    \label{fig:scenario_pCO2_all}
\end{figure}

We use geological data to quantify the changes in the area of I\&NG over the past 20~m.y. (described in detail in \textit{Paleoshoreline Reconstruction and Geological Synthesis}). To generate the lithologic map used in the ``removed I\&NG'' scenario (Fig. \ref{fig:ING_fracs}), we remove all land associated with arc-continent collision in I\&NG. To generate the lithologic map used in the ``reduced I\&NG + Sunda Shelf'' scenarios (Fig. \ref{fig:ING_fracs}), we take the estimated paleoshorelines of I\&NG for 15, 10, and 5~Ma, then remove all land that falls outside of these bounds. The slope and lithologic classification of each pixel is identical to that of the present day, provided that these pixels fall within the estimated paleoshorelines of I\&NG for their respective time slice. This means that while we are changing the amount of emergent land, we are not changing the latitudinal or longitudinal position and are keeping the slope of emerged pixels the same. Given that similar distributions of lithology and slope are reasonable in the past and that latitudinal translation has been relatively minor, we consider this simplification to be reasonable to first-order (Fig. XXX). We then add land that is not exposed today (i.e. the Sunda Shelf). However, the assignment of lithology, runoff, and slope for pixels on the Sunda Shelf is not trivial, given that the shelf is not currently exposed. As discussed in \textit{Paleoshoreline Reconstruction and Geological Synthesis}, the Sunda Shelf is a flat-lying, relatively stable platform of continental crust. Islands between Malaysia/Sumatra and Borneo are composed of granite and are surrounded by shallow marine siliciclastic sediments \citep{Darmadi2007a, Hall2009a, Hall2013b}. However, basement highs do not appear in the Gulf of Thailand, and seismic and drill core data suggest siliciclastic fluvial systems draining out toward the South China Sea when the shelf was exposed \citep{Darmadi2007a}. We therefore assign pixels of the Sunda Shelf between Malaysia/Sumatra and Borneo to be 50\% felsic and 50\% sediment, and pixels further to the north to be 100\% sediment. For the slope, we take the mean slope value (0.0038~m/m) of relatively flat land on eastern Sumatra, and assign this value to all pixels of the Sunda Shelf.  For the runoff, we linearly interpolate along latitude bands between present-day land pixels to obtain runoff values for the presently-submerged Sunda Shelf. This approach preserves large N-S variation in runoff associated with the large-scale Hadley circulation and therefore should be reasonable to first-order, although it does not accurately capture the smaller-scale interactions between land and ocean/atmosphere circulation.
